Library Learning in Neurally-Guided Bayesian Program Induction

### 2018-10-26 nips_2018.tex
almost, ready

Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during exploration, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference,stuhlmuller2013learning}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
%% \noindent\\\textbf{Contributions.} 
(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compression):
\\\noindent \textbf{Explore: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

%% Learning the DSL eases the difficulty of synthesis
%% by exposing a domain-specific basis for constructing programs.
%% Another complementary means of meeting search is to learn a
%% bottom-up \emph{recognition model}

\subsection{Exploration: Searching for Programs}\label{explorationSection}

@@

\subsection{Compilation: Learning a Neural Recognition Model}\label{recognitionSection}

@@

How should we get the data to train $q$?
This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by search.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
 equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
 where the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
 The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
\begin{comment}
  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\subsection{Compression: Learning a Generative Model (a DSL)}\label{grammarInductionSection}

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Tbl.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative,hwang2011inducing}.
Care must be taken
to ensure that
this `fragmenting' obeys variable scoping rules;
Section 4 of the supplement gives an overview of Fragment Grammars
and how we adapt them to the lexical scoping rules of $\lambda$-calculus.
\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{0.2\textwidth}
  \begin{tikzpicture}[scale=0.7]
    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
    %%     \footnotesize Program:\\
    %%     \code{($\lambda$ (x) (+ (- x) 1))}
    %% \end{tabular}};
    %\node at (,0.5) {\code{cons}};
    \node [rotate=90] at (-2.3,-0.5) {\small program};

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
highlighted in {\orange{orange}},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
\end{figure}
\begin{figure}%{R}{0.6\textwidth}\vspace{-1cm}
\centering  \begin{minipage}{0.9\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hfill(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\footnotesize{\textbf{Compile}})
        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hfill\textbf{(Explore)}}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\footnotesize{\hfill\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}
To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

>>>>

Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during the Explore step, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference,stuhlmuller2013learning}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
%% \noindent\\\textbf{Contributions.} 
(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:

\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compress):
\\\noindent \textbf{Explore: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Compress: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the Compile step: $q$ is trained to approximate the
posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

\textbf{Compile: learning to tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.
We train $q$ both on programs found during the Explore step and
on samples from the current DSL.

\begin{wrapfigure}{r}{0.6\textwidth}\vspace{-0.55cm}\centering
  \begin{tikzpicture}[scale=0.9]
    {\footnotesize
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compress)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compile)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compile)}} (recognitionModel.west);

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
  \end{scope}}
    \end{tikzpicture}
  \caption{How these steps bootstrap each other.}  \label{feeding}\vspace{-1cm}
\end{wrapfigure}%
Crucially, each of these three steps bootstraps the others (Fig.~\ref{feeding}): improving either the DSL or the recognition model
makes search easier, so we find more programs solving tasks;
both improving the DSL and solving more tasks expands the training data for the recognition model;
and finding more programs that solve tasks
gives more data from which to learn a DSL.

%% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

%% Learning the DSL eases the difficulty of synthesis
%% by exposing a domain-specific basis for constructing programs.
%% Another complementary means of meeting search is to learn a
%% bottom-up \emph{recognition model}
\vspace{0.5cm}Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
For added robustness,
we interleave an extra program search step (Explore) before training the recognition model,
and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.

\begin{figure}[h]
\centering  \begin{minipage}{0.9\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \hfill(\textbf{Explore})
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\textbf{Compile})
        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \hfill\textbf{(Explore)}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\hfill\textbf{(Compress)}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}

\subsection{Explore: Searching for Programs}\label{explorationSection}

@@

\subsection{Compile: Learning a Neural Recognition Model}\label{recognitionSection}

@@

How should we get the data to train $q$?
This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how the wake-sleep algorithm for the Helmholtz machine
trains its recognition model during its sleep phase~\cite{dayan1995helmholtz}.
The advantage of training on samples, or ``fantasies,'' is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
to train $q$ on the (program, task)
pairs discovered by the Explore step.
The advantage here
is that the training data is much higher quality,
because we are training on real tasks.
Due to these complementary advantages,
we train on both these sources of data.

Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
 equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
 where the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks trains $q$ on the real data; taking it over samples from the generative model trains $q$ on ``fantasies.''
 The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the fantasy ($\mathcal{L}_{\text{f}}$) and real-data ($\mathcal{L}_{\text{r}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{r} + \mathcal{L}_\text{f}$:
\begin{align}\nonumber
\mathcal{L}_{\text{f}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{r}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
\begin{comment}
  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

\subsection{Compress: Learning a Generative Model (a DSL)}\label{grammarInductionSection}

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Tbl.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative,hwang2011inducing}.
Care must be taken
to ensure that
this `fragmenting' obeys variable scoping rules;
Section 4 of the supplement gives an overview of Fragment Grammars
and how we adapt them to the lexical scoping rules of $\lambda$-calculus.
\begin{figure}[h!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{0.2\textwidth}
  \begin{tikzpicture}[scale=0.7]
    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
    %%     \footnotesize Program:\\
    %%     \code{($\lambda$ (x) (+ (- x) 1))}
    %% \end{tabular}};
    %\node at (,0.5) {\code{cons}};
    \node [rotate=90] at (-2.3,-0.5) {\small program};

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
highlighted in {\orange{orange}},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
\end{figure}
To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
